---
title:  "IDS 2021. Problem Set 2"
author: "Muhammad Zahidul Islam Miaji"
date:   "Last updated at `r Sys.time()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-data, message = FALSE}
#install.packages("gapminder")
library(gapminder)
library(tidyverse)

gapminder07 <- filter(gapminder, year %in% 2007)
```

## Q1 

How many variables and how many observations are in the original gapminder data? How about for the data subset for 2007?

```{r}
glimpse(gapminder)
observ_gapminder <- gapminder
observ_gapminder
gapminder
gapminder07 <- filter(gapminder, year %in% 2007)
```

Total observation for "gapminder" is 1704 of 6 variables. And year of 2007 (subset 2007) is 142 observation of 6 variables


## Q2 

Let's create a new variable `real_gdp` by multiplying the `gdpPercap` variable with the `pop` variable (hint: 1. GDP per capita is calculated by dividing the real GDP by population, 2. use the `mutate` function to create the new variable)

```{r}
gapminder_realGdp <- gapminder %>%
  mutate(real_gdp = gdpPercap * pop)
gapminder_realGdp

```

## Q3 

Next, let's compute the average life expectancy by continent in the year 2007. (hint: 1. so you'll need to use the `gapminder07` data. 2. you wilcontinl probably need to use functions such as `filter`, `group_by`, `summarize`, and `mean`, 3. make sure to use `na.rm=TRUE` for your mean function to avoid observations from dropping out of your data inadvertantly).

```{r}
avg_life <- gapminder07 %>%
  filter(!is.na(lifeExp), !is.na(continent)) %>%
  group_by(lifeExp, continent) %>%
  summarise(avg_LE = round(mean(lifeExp, na.rm = TRUE)))
avg_life            
```



## Q4 

Next, let's compute the average life expectancy by continent and year from the full dataset (`gapminder`). Draw a line plot over time to examine the trend.

```{r}
gapminder %>%
    group_by(continent, year) %>%
    summarise(lifeExp=mean(lifeExp)) %>%
    ggplot(aes(x=year, y=lifeExp, color=continent)) +
     geom_line(size=1) + 
     geom_point(size=1.5)
    

```

## Q5 

```{r}
gapminder%>%
  filter (year %in% c(1957, 1977, 1997))%>%
  ggplot(aes(x = log(gdpPercap), color= as.factor(year), fill= as.factor(year))) +
  geom_density(alpha= 0.2)
  
```


## Q6. 

Create a table that shows the Life Expectancy in 2007 for the countries in the Americas. Report only the country and the life expectancy variables in the table. (hint: you will probably need to use functions such as `filter`, `select`, `tableGrob`, `grid.arrange`, or `kable`).

```{r}
library(gridExtra)
```
```{r, fig.height=8, fig.width=4}
gapminder %>%
  filter(year== 2007, continent== "Americas")%>%
  select(country, lifeExp)%>%
  tableGrob(cols = c("country", "lifeExp")) %>%
  grid.arrange()

```
